---
title: "Statistical Calculation and Software"
author: "Assignment 1"
date: "Hanbin Liu 11912410"
output:
  #html_document
  pdf_document:
    keep_tex: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
setwd("C:/Users/HP/Desktop/R_project/R_assignments")
library(ggplot2)
library(dplyr)
library(hrbrthemes)
```

## 1.1
### (a)
```{r}
data('iris')
```

```{r}
Species <- factor(iris$Species, labels = c("a", "b", "c"))
iris <- data.frame(iris[, 1:4], Species)
table(iris$Species)
```

### (b)
```{r collapse=TRUE}
ggplot(iris, aes(x = Species, y = Sepal.Length)) +
  geom_boxplot(fill = "darkgreen") +
  ggtitle("Sepal length distribution for each type of irises") +
  theme(plot.title = element_text(hjust = 0.5),
        text = element_text(size = 14, family = "serif"))
```

### (c)
```{r}
ind <- which(iris["Sepal.Length"] > 5.5)
iris_ed <- iris[ind,]

length <- iris_ed$Petal.Length
width <- iris_ed$Petal.Width
value <- c(length, width)
group <-
  c(rep('Petal.Length', length(width)), rep('Petal.width', length(width)))
species <- c(iris_ed$Species, iris_ed$Species)
data <- data.frame(species, group, value)

ggplot(data, aes(x = species, y = value, fill = group)) +
  geom_boxplot() + xlab("Species") + ylab("Value") +
  theme(legend.position = "top",
        text = element_text(size = 14, family = "serif"))
```

## 1.2
```{r}
F2000 <- read.csv("F2000.csv")
```

### (a)&(b)
```{r}
skewness <- function(x) {
  xixi <- x - mean(x)
  skewness <- sum(xixi ^ 3) / ((length(x) - 1) * var(x) ^ 1.5)
  return(skewness)
}

x <- F2000[, "marketvalue"]
skewness(x)
skewness(log(x))
skewness(1 - x ^ -1)
skewness((x ^ 0.25 - 1) / 0.25)
```

### (c)
```{r fig.height=10, fig.width=20, warning=FALSE}
library(patchwork)
p1 <- ggplot(F2000, aes(x = marketvalue)) +
  geom_histogram(binwidth = 10,
                 fill = "#000000",
                 color = "#e9ecef") +
  ggtitle("Histogram of marketvalue") +
  theme(plot.title = element_text(hjust = 0.5),
        text = element_text(size = 14, family = "serif")) +
  xlab(expression(italic(x)))

p2 <- ggplot(F2000, aes(x = log(marketvalue))) +
  geom_histogram(
    binwidth = 0.5,
    fill = "#3E432E",
    color = "#e9ecef",
    alpha = 0.8
  ) +
  ggtitle("Histogram of log-transformed marketvalue") +
  theme(plot.title = element_text(hjust = 0.5),
        text = element_text(size = 14, family = "serif")) +
  xlab(expression(italic(log(x))))

p3 <- ggplot(F2000, aes(x = 1 - (marketvalue) ^ -1)) +
  geom_histogram(binwidth = 0.1,
                 fill = "#616F69",
                 color = "#e9ecef") +
  ggtitle("Histogram of inverse-transformed marketvalue") +
  theme(plot.title = element_text(hjust = 0.5),
        text = element_text(size = 14, family = "serif")) +
  xlim(c(-10, 1)) +
  xlab(expression(italic(frac(x ^ -1 - 1,-1))))

p4 <- ggplot(F2000, aes(x = marketvalue ^ 0.25 - 1)) +
  geom_histogram(binwidth = 0.1,
                 fill = "#A7D129",
                 color = "#e9ecef") +
  ggtitle(expression(paste('Histogram with ', italic(lambda), " = 0.25"))) +
  theme(plot.title = element_text(hjust = 0.5),
        text = element_text(size = 14, family = "serif")) +
  xlab(expression(italic(frac(x ^ 0.25 - 1, 0.25))))

p1 + p2 + p3 + p4
```

### (d)
```{r}
ind <- which(is.na(F2000[, "profits"]))
F2000[, "name"][ind]
fivenum(F2000[, "sales"][ind])
```

### (e)
```{r}
# number of different countries
length(table(F2000$country))

xixi <- data.frame(table(F2000$country))
mean <- 1:61
median <- mean
for (i in 1:61) {
  ind <- which(F2000$country == xixi[i, 1])
  mean[i] <- mean(F2000[ind,]$assets)
  median[i] <- median(F2000[ind,]$assets)
}

countries <-
  data.frame(
    country = xixi[, 1],
    num_of_companies = xixi[, 2],
    mean_assets = mean,
    median_assets = median
  )
write.table(countries, file = "countries.txt")
```

### (f)
```{r}
# 2 ways
way_1 <- F2000[F2000$sales > 100,]
way_2 <- F2000[which(F2000$sales > 100),]
# sort
selected_data <- way_1[-c(3, 4, 8)]
ind_1 <- order(selected_data$sales, decreasing = TRUE)
selected_data[ind_1,]
ind_2 <- order(selected_data$assets)
selected_data[ind_2,]
# plot
ggplot(selected_data, aes(x = assets, y = profits)) +
  geom_point(color = "blue") +
  ggtitle("Scatter plot") +
  theme(plot.title = element_text(hjust = 0.5),
        text = element_text(size = 14, family = "serif")) +
  xlab("Assets") +
  ylab("Profits")
```

### (g)
- method 1(my function)
```{r}
myKNN <-
  function(data, train_col, pred_col, k, scale, M) {
    #scale: scale() or min-max; M: mean or median
    X <- data[, train_col]
    if (scale == "min-max") {
      # min-max normalization
      min <- apply(X, 2, min)
      max <- apply(X, 2, max)
      MIN <- matrix(rep(min, nrow(X)), ncol = length(min), byrow = T)
      MAX <- matrix(rep(max, nrow(X)), ncol = length(max), byrow = T)
      X <- (X - MIN) / (MAX - MIN)
    } else if (scale == "scale") {
      X <- scale(X)
    }
    # find index of missing value
    ind <- which(is.na(data[, pred_col]))
    # index of training samples
    train_id <- setdiff(1:nrow(X), ind)
    # operation w.r.t sample p
    for (p in ind) {
      dist <- replicate(nrow(X),-1)
      for (i in train_id) {
        temp <- 0
        for (j in 1:length(train_col)) {
          temp <- temp + (X[i, train_col[j]] - X[p, train_col[j]]) ^ 2
        }
        dist[i] <- sqrt(temp)
      }
      # select k neighbors w.r.t sample p
      selid <- which(dist %in% sort(dist[dist >= 0])[1:k])
      # mean of k neighbors on pred_col    #or median or mode
      if (M == "mean") {
        data[p, pred_col] <- mean(data[selid, pred_col])
      } else if (M == "median") {
        data[p, pred_col] <- median(data[selid, pred_col])
      }
    }
    # show
    result <- data.frame(data[ind,-c(4, 5, 7, 8)])
    return(result)
  }
```
```{r}
myKNN(F2000,
      c("sales", "assets", "marketvalue"),
      "profits",
      10,
      "scale",
      "mean")
myKNN(F2000,
      c("sales", "assets", "marketvalue"),
      "profits",
      10,
      "min-max",
      "mean")
myKNN(F2000,
      c("sales", "assets", "marketvalue"),
      "profits",
      10,
      "scale",
      "median")
myKNN(F2000,
      c("sales", "assets", "marketvalue"),
      "profits",
      10,
      "min-max",
      "median")
```

- method 2(use package)
```{r warning=FALSE}
library(caret)
library(RANN)
library(lattice)
# save mean and sd for de-scaling
train_col <-
  c("sales", "assets", "marketvalue")
pred_col <- "profits"
ind <- which(is.na(F2000[, pred_col]))
mean <- mean(F2000[setdiff(1:nrow(F2000), ind), pred_col])
sd <- sd(F2000[setdiff(1:nrow(F2000), ind), pred_col])
# knn imputation
model <-
  preProcess(F2000[, c(train_col, pred_col)], method = "knnImpute", k = 10)
prediction <- predict(model, F2000)
# de-scaling
prediction[, pred_col] <- mean + sd * prediction[, pred_col]
# show
result <- data.frame(prediction[ind,-c(4, 5, 7, 8)])
result
```

## 1.3
### (a)
```{r}
ggplot(data.frame(x = c(0, 8)), aes(x = x)) +
  ggtitle("Density of Chi-squared distribution") +
  stat_function(fun = dchisq,
                args = list(df = 4),
                aes(colour = "freedom = 4")) +
  stat_function(fun = dchisq,
                args = list(df = 6),
                aes(colour = "freedom = 6")) +
  scale_colour_brewer(palette = "Accent") +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.position = "top",
    text = element_text(size = 14, family = "serif")
  ) +
  labs(colour = "") +
  xlab(expression(italic(x))) +
  ylab(expression(italic(f(x))))
```
```{r}
s1 <- pchisq(7, df = 5) - pchisq(1, df = 5)
s1
s2 <- 1 - pchisq(3, df = 5)
s2
```

### (b)
The cdf of $X$ is given by 
\begin{equation*}
F(x) = \int_0^x \lambda e^{-\lambda t} dt = 1 - e^{-\lambda x}.
\end{equation*}
Then, $F(1) = 1-e^{-\lambda},$ $F(6) = 1 - e^{-6\lambda}.$
```{r}
lambda <- -log(0.8)
F6 <- 1 - 0.8 ^ 6
sprintf('lambda:   %f', lambda)
sprintf('Pr(X<=6): %f', F6)
```

### (c)
$X_i \sim U(0,1).$ Let $Y = X_1+2X_2, Z = X_1,$ then$\frac{\partial(x_1,x_2)}{\partial(y,z)}$ = 
$$
\begin{vmatrix}
0 & 1\\
\frac{1}{2} & \frac{-1}{2} 
\end{vmatrix}$$
= $\frac{-1}{2}.$
Then, $f_{Y,Z}(y,z) = \int_0^1 \int_0^1  f_{X_1,X_2}(x_1,x_2)\cdot \frac{1}{2}dx_1dx_2 = \frac{1}{2}, \quad z\leq y \leq z+2,\ 0\leq z \leq 1$
Hence, if $y<0,$ then $f_Y(y) = 0;$ if $0\leq y < 1,$ then$f_Y(y) = \int_0^y \frac{1}{2}dz = \frac{y}{2};$ if $1 \leq y < 2,$ then$f_Y(y) = \int_0^1 \frac{1}{2} dz = \frac{1}{2};$ if $ 2\leq y <3,$ then$f_Y(y) = \int_{y-2}^1 \frac{1}{2} dz = \frac{3-y}{2};$ if$y \geq 3,$ then$f_Y(y) = 0.$ Therefore, the pdf of $Y$ is given by
\begin{equation}
f_Y(y)=
\begin{cases}
0           & y < 0, \\
\frac{y}{2} & 0 \leq y < 1, \\
\frac{1}{2} & 1 \leq y < 2,\\
\frac{3-y}{2} & 2 \leq y < 3, \\
0           & y \geq 3.
\end{cases}
\end{equation}
```{r}
x <- c(-1, 0, 1, 2, 3, 4)
y <- c(0, 0, 0.5, 0.5, 0, 0)
ggplot(data.frame(x, y), aes(x = x, y = y)) +
  geom_line(colour = 'thistle') +
  ggtitle(expression(paste("Density of ", italic(Y)))) +
  theme(plot.title = element_text(hjust = 0.5),
        text = element_text(size = 14, family = "serif")) +
  ylab(expression(italic(f(y)))) +
  xlab(expression(italic(y)))
```

### (d)
```{r}
ggplot(data.frame(x = c(-4, 4)), aes(x = x)) +
  ggtitle(expression(
    paste(
      "Density of standard normal distribution and ",
      italic(t),
      "-distribution"
    )
  )) +
  stat_function(fun = dnorm,
                args = list(0, 1),
                aes(colour = "standard normal")) +
  stat_function(fun = dt,
                args = list(1),
                aes(colour = "t(1) distribution")) +
  scale_colour_brewer(palette = "Accent") +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.position = "top",
    text = element_text(size = 14, family = "serif")
  ) +
  labs(colour = "") +
  xlab(expression(italic(x))) +
  ylab(expression(italic(f(x))))
```

$t(1)$ distribution's tail is heavy.

### (e)
Note that
\begin{equation*}
  1 = \sum_{x=0}^{{\rm min}\{m,n\}} {\rm Pr}(X=x) = \sum_{x=0}^{{\rm min}\{m,n\}} \frac{\binom{m}{x} \binom{N-m}{n-x}}{\binom{N}{n}}.
\end{equation*}
That is,
\begin{equation*}
  \sum_{x=0}^{{\rm min}\{m,n\}} \binom{m}{x} \binom{N-m}{n-x} = \binom{N}{n}.
\end{equation*}
The expectation of $X$ is given by 
\begin{equation*}
  E(X) = \sum_{x=0}^{{\rm min}\{m,n\}} x{\rm Pr}(X=x) = \frac{1}{\binom{N}{n}} \sum_{x=1}^{{\rm min}\{m,n\}} x \binom{m}{x} \binom{N-m}{n-x}.
\end{equation*}
Using $x\binom{m}{x} = m\binom{m-1}{x-1},$ we have 
\begin{equation*}
  \begin{aligned}
    E(X) &= \frac{1}{\binom{N}{n}} \sum_{t=0}^{{\rm min}\{m-1,n-1\}} m\binom{m-1}{t} \binom{(N-1)-(m-1)}{(n-1)-t} \quad (t = x-1)\\
         &= \frac{m}{\binom{N}{n}} \sum_{t=0}^{{\rm min}\{m-1,n-1\}} \binom{m-1}{t} \binom{(N-1)-(m-1)}{(n-1)-t}\\
         &= \frac{m}{\binom{N}{n}} \binom{N-1}{n-1}\\
         &= \frac{mn}{N}.
  \end{aligned}
\end{equation*}
Similarly, using $x^2 = x(x-1) + x$ and the same skill, we can obtain that $E(X^2) = \frac{mn}{N} + \frac{mn(m-1)(n-1)}{N(N-1)}$. Then 
\begin{equation*}
  {\rm Var}(X) = E(X^2) - (EX)^2 = \frac{mn(N-m)(N-n)}{N^2(N-1)}.
\end{equation*}
```{r}
p <- replicate(9, 0)
name <- p
E <- 0
E2 <- 0
for (k in 0:8) {
  name[k + 1] <- paste("Pr(X = ", k, ")", sep = "")
  p[k + 1] <- (choose(28, k) * choose(17, 8 - k)) / choose(45, 8)
  E <- E + k * p[k + 1]
  E2 <- E2 + k ^ 2 * p[k + 1]
}
table <- data.frame(name, p)
table
Var <- E2 - E ^ 2
paste("Expectation:", E)
paste("Variation:  ", Var)
```

### (f)
${\rm Pr}(X > 8) \leq 0.4$ is equivalent to $\sum_{k=0}^8{\rm Pr}(X = k) \geq 0.6.$
```{r}
f <- function(lambda) {
  sum(dpois(0:8, lambda)) - 0.6
}
if (f(1) > 0 && f(10) < 0) {
  lambda <- uniroot(f, lower = 1, upper = 10)[[1]]
  lambda
}
```

### (g)
```{r}
x <- seq(0, 5, 0.1)
y <- dgamma(x, 1, 3)
ggplot(data.frame(x, y), aes(x = x, y = y)) +
  geom_point(color = 'purple') +
  stat_function(fun = dexp,
                args = list(3),
                aes(colour = "Exponential(3)")) +
  ggtitle("Density of Exponential(3) and Gamma(1, 3) distributions") +
  scale_colour_brewer(palette = "Accent") +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.position = "top",
    text = element_text(size = 14, family = "serif")
  ) +
  labs(colour = "") +
  xlab(expression(italic(x))) +
  ylab(expression(italic(f(x)))) 
# Purple points are generated from Gamma(1, 3) distribution
```

### (h)
```{r}
prob <-
  pretty(0:1, 20)
prob[21] <-
  1  # 1 - prob[21] = -2.220446e-16 without this assignment
size <- replicate(21, 0)
for (i in 1:21) {
  n <- 1
  p <- 1
  while (p > 1 - prob[i]) {
    n <- n + 1
    p <- p * (365 - n + 1) / 365
  }
  size[i] <- n
}
result <- data.frame(size, prob)
result
```

### (i)
- method 1 (visualization)
```{r}
myplot <- function(n, prob) {
  times <- 10000
  x <- rbinom(times, n, prob)
  z <- (x - n * prob) / (sqrt(n * prob * (1 - prob)))
  p <- ggplot(data.frame(value = z), aes(x = value)) +
    stat_function(fun = dnorm, aes(colour = 'Standard normal distribution')) +
    geom_bar(aes(y = ..count.. / (sqrt(2 * pi) * max(..count..)) , colour = 'Zn')) +
    scale_colour_brewer(palette = "Accent") +
    theme(
      plot.title = element_text(hjust = 0.5),
      legend.position = "top",
      text = element_text(size = 14, family = "serif")
    ) +
    labs(colour = "") +
    xlab(expression(italic(x))) +
    ylab(expression(italic(f(x))))
  return(p)
}
```
```{r}
prob <- runif(1, min = 0, max = 1)
myplot(10, prob)
myplot(100, prob)
```

- method 2 (calculate) - a and b not fixed
```{r}
# simulation times of Zn
times <- 5000
# randomly choose p
p <- runif(1, min = 0, max = 1)
# calculate Pr(a < Zn <= b) for n = 10, 100, 1000, 10000
simulation <-
  rep(0, 4)
true <- rep(0, 4)
a <- rep(0, 4)
b <- rep(0, 4)
for (n in c(10, 100, 1000, 10000)) {
  x <- rbinom(times, n, p)
  z <- (x - n * p) / (sqrt(n * p * (1 - p)))
  # randomly choose a and b among the range of z
  nums <- runif(2, min = min(z), max = max(z))
  a[log10(n)] <- min(nums)
  b[log10(n)] <- max(nums)
  simulation[log10(n)] <-
    length(z[z > a[log10(n)] & z <= b[log10(n)]]) / length(z)
  # value of integral
  true[log10(n)] <- pnorm(b[log10(n)]) - pnorm(a[log10(n)])
}
# show
n <- c(10, 100, 1000, 10000)
comparision <- data.frame(n, a, b, simulation, true)
comparision
```

-- a and b fixed
```{r}
a <- -1
b <- 1
times <- 5000
p <- runif(1, min = 0, max = 1)
simulation <- rep(0, 4)
true <- rep(0, 4)
for (n in c(10, 100, 1000, 10000)) {
  x <- rbinom(times, n, p)
  z <- (x - n * p) / (sqrt(n * p * (1 - p)))
  simulation[log10(n)] <- length(z[z > a & z <= b]) / length(z)
  true[log10(n)] <- pnorm(b) - pnorm(a)
}
# show
n <- c(10, 100, 1000, 10000)
comparision <- data.frame(n, a, b, simulation, true)
comparision
```

### (j)
- method 1 
```{r}
# lambda: lambda; times: times of the generation of Sn for each n; K: k = 1,...,K
poissonlimit <- function(lambda, times, K) {
  simulation <- matrix(rep(0, 5 * K), 5, K)
  
  for (n in c(10, 100, 1000, 10000)) {
    sn <- rbinom(times, n, lambda / n)
    for (k in 1:K) {
      simulation[log10(n), k] <- length(sn[sn == k]) / length(sn)
    }
  }
  # show
  result <- data.frame(c(10, 100, 1000, 10000, "true"), simulation)
  colnames(result) <- c("n\\k", 1:9)
  result[5, 2:10] <- dpois(1:9, lambda)
  return(result)
}
```
```{r}
poissonlimit(3, 5000, 9)
poissonlimit(5, 5000, 9)
poissonlimit(7, 5000, 9)
poissonlimit(9, 5000, 9)
```

- method 2 (visualization)
```{r}
pois_binom <- function(lambda, n) {
  p <- lambda / n
  x <- 1:20
  y_pois <- dpois(x, lambda)
  y_binom <- dbinom(x, n, p)
  pp <- ggplot(data.frame(x, y_pois, y_binom)) +
    geom_bar(aes(x = x, y = y_pois), stat = 'identity') +
    geom_point(aes(x = x, y = y_binom), col = '#A4D65E', size = 2) +
    ggtitle("Bar for Poission & Points for Binomial") +
    theme(
      plot.title = element_text(hjust = 0.5),
      legend.position = "top",
      text = element_text(size = 14, family = "serif")
    ) +
    xlab(expression(italic(x))) +
    ylab(expression(italic(p(x))))
  return(pp)
}
```
```{r}
pois_binom(3, 10)
pois_binom(3, 100)
pois_binom(3, 1000)
##
pois_binom(9, 1000)
```

### (k)
Note that in $R,$ the geometric distribution with $prob = p$ has density 
\begin{equation*}
p(x) = p(1-p)^x,
\end{equation*}
for $x = 0, 1, 2, 3, ...,$ which is different from what we tipically learned in class: 
\begin{equation*}
p(x) = p(1-p)^{x-1},
\end{equation*}
for $x = 1,2,3,....$ Therefore, we should use $pgeom(m + n + 1)$ instead of $pgeom(m + n)$ for geometric distribution.

- method 1 (cdf method)
```{r}
# p: p or lambda; type: geometric or poisson
memoryless_1 <- function(m, n, p, type) {
  if (type == 'Geometric') {
    if (round((1 - pgeom(m + n + 1, p)), 4) == round(((1 - pgeom(m, p)) *
                                                      (1 - pgeom(n, p))), 4)) {
      print('Geometric  distribution has memoryless  property')
    } else{
      print('Geometric  distribution does not have memoryless  property')
    }
  } else if (type == 'Poisson') {
    if (round((1 - ppois(m + n, p)), 4) == round(((1 - ppois(m, p)) *
                                                  (1 - ppois(n, p))), 4)) {
      print('Poisson  distribution has memoryless  property')
    } else{
      print('Poisson  distribution does not have memoryless  property')
    }
  }
}
```
```{r}
memoryless_1(5, 7, 0.2, 'Geometric')
memoryless_1(11, 103, 0.4, 'Geometric')
memoryless_1(75, 49, 0.6, 'Geometric')
memoryless_1(5, 7, 3, 'Poisson')
memoryless_1(11, 13, 7, 'Poisson')
memoryless_1(9, 40, 21, 'Poisson')
```

- method 2 (simulation method)
```{r}
# p: p or lambda; T: simulation times; k: number of x per time; type: geometric or poisson
memoryless_2 <- function(p, T, k, type) {
  result <- matrix(rep(0, 4 * T), T, 4)
  for (t in 1:T) {
    if (type == "geometric") {
      x <- rgeom(k, p)
    } else if (type == "Poisson") {
      x <- rpois(k, p)
    }
    # in order to get a good simulation, m and n are selected as follows
    m <-
      round(runif(1, min = mean(x) - sd(x), max = mean(x) + sd(x)))
    n <-
      round(runif(1, min = mean(x) - sd(x), max = mean(x) + sd(x)))
    if (type == "geometric") {
      p_condition <- length(x[x > m + n + 1]) / length(x[x > m])
    } else if (type == "Poisson") {
      p_condition <- length(x[x > m + n]) / length(x[x > m])
    }
    p2 <- length(x[x > n]) / length(x)
    result[t,] <- c(p_condition, p2, m, n)
  }
  result <- data.frame(result)
  colnames(result) <- c("Pr(X > m+n|X > m)", "Pr(X > n)", "m", "n")
  return(result)
}
```
```{r}
memoryless_2(0.5, 10, 10000, "geometric")
memoryless_2(0.1, 10, 10000, "geometric")
memoryless_2(0.01, 10, 10000, "geometric")
```

From the result, we can see that geometric distribution has memoryless property. 
```{r}
memoryless_2(1, 10, 10000, "Poisson")
memoryless_2(5, 10, 10000, "Poisson")
memoryless_2(10, 10, 10000, "Poisson")
memoryless_2(50, 10, 10000, "Poisson")
```
From the result, Poisson distribution does not have memoryless property.



















